---
title: "P8105 Homework 1"
author: "dxy2101"
date: "September 24th, 2022"
output: github_document
---

This assignment focuses on getting started with R, best practices, and working in Github.

```{r setup1}
library(tidyverse)
```

```{r setup2}
install.packages("palmerpenguins", repos = "http://cran.us.r-project.org")
```

# Problem 1

### Load and describe `penguins` dataset.

```{r penguins}
data("penguins", package = "palmerpenguins")

# List variables in dataset
names(penguins)

nrow(penguins) # Number of observations
ncol(penguins) # Number of variables

# Calculate mean flipper length
mean(penguins$flipper_length_mm, na.rm=TRUE)
```

The `penguins` dataset includes 8 variables describing a sample of 344 individual penguins. This dataset may be useful for understanding the different compositions of different penguin species across e.g. geographies, years, sex, and among the physical attributes themselves. 

The following variables are included:
* species
* island of origin
* bill length (mm)
* bill depth (mm)
* flipper length (mm)
* body mass (g)
* sex
* year data was collected

From `penguins`, we found that the mean flipper length, excluding missing values, is 200.9152 mm.

### Create scatterplot for flipper length and bill length

```{r scatterplot}
# Create a dataframe containing variables for the plot
plot_df = tibble(
  x = penguins$bill_length_mm,
  y = penguins$flipper_length_mm
)

# Create scatter plot with colors indicating species type
Species = penguins$species

ggplot(plot_df, aes(x=x, y=y, color=Species)) + geom_point()

# Export scatter plot
ggsave("bill_flipper_plot.pdf", height = 4, width = 6)
```

# Problem 2

### Create a new data frame.

```{r new df}
prob2_df = tibble(
  vec_numeric = rnorm(10), # random sample of 10 N(0,1)
  
  vec_logical = vec_numeric > 0, # indicate whether elements of sample > 0
  
  vec_character = c("How", "are", "you", "doing", "I", "hope", "you",
                    "are", "doing", "well"), # character vector of length 10
  
  vec_factor = factor(c("yellow", "red", "blue", "yellow", "red", "blue",
                        "yellow", "red", "blue", "yellow")) # factor with 3 levels + length 10
)
```

Try to take the mean of each variable.

```{r mean}
mean(pull(prob2_df, vec_numeric)) # take mean of numeric vector
mean(pull(prob2_df, vec_logical)) # take mean of logical vector
mean(pull(prob2_df, vec_character)) # take mean of character vector
mean(pull(prob2_df, vec_factor)) # take mean of factor vector
```

Taking the mean of `vec_numeric` and `vec_logical` return a value. Taking the mean of `vec_character` and `vec_factor` is non-applicable.

### Convert all variables to numeric.

```{r as.numeric, eval = FALSE}
as.numeric(pull(prob2_df, vec_logical)) # convert vec_logical to numeric
as.numeric(pull(prob2_df, vec_character)) # convert vec_character to numeric
as.numeric(pull(prob2_df, vec_factor)) # convert vec_factor to numeric
```

Converting `vec_logical` to numeric resulted in a binary numeric output (i.e. 1, 0). 1 indicating TRUE and 0 indicating FALSE. Converting `vec_character` to numeric resulted in outputs that were also listed as NA. Converting `vec_factor`to numeric resulted in three types of numeric outputs (i.e. 1, 2, 3) to match the three different levels.

The binary nature of `vec_logical` explains the numeric output after taking its mean. However, this does not explain why the same could not happen for `vec_factor`.